---
title: "Results_Final"
author: "Kamerin Sandlin"
date: "2026-04-14"
output: 
    github_document: default
    html_document: default
---

### We identified soybean mosaic virus, tomato spotted wilt virus, peanut muttle virus, and pea enation mosaic virus -1 in the seven seoybean samples analyzed. The following are depictions and more data on these results. 

# NOTE
### This is the progress made to the point of April. We are still working to validate tomato spotted wilt virus in soybean 1, 2-3, and 6. 

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

# Data
data <- data.frame(
    Sample = c("SB1", "SB2-3", "SB5", "SB6", "SB7-8", "SB9", "SB10"),
    Field = c("Old Rotation", "Old Rotation", "Cullers", "Old Rotation", "Old Rotation", "Cullers", "Cullers"),
    SMV = c(1,1,0,1,1,1,1),
    TSWV = c(0,0,0,0,1,1,0),
    PEMV = c(1,0,0,0,0,0,0),
    PMV = c(0,0,0,1,0,0,0)
)

# Convert to long format
data_long <- data %>%
    pivot_longer(cols = SMV:PMV, names_to = "Virus", values_to = "Presence")

# Order samples
data_long$Sample <- factor(data_long$Sample,
                           levels = c("SB1", "SB2-3", "SB5", "SB6", "SB7-8", "SB9", "SB10"))

# Plot
ggplot(data_long, aes(x = Virus, y = Sample, fill = factor(Presence))) +
    geom_tile(color = "black") +
    
    scale_fill_manual(
        name = "Virus Detection",
        values = c("0" = "white", "1" = "blue"),
        labels = c("0" = "No detection", "1" = "Virus detected")
    ) +
    
    theme_minimal() +
    theme(
        axis.text.y = element_text(
            color = ifelse(data$Field[match(levels(data_long$Sample), data$Sample)] == "Old Rotation",
                           "black", "red"),
            size = 12
        ),
        axis.title.y = element_blank()
    ) +
    
    labs(
        x = "Virus",
        y = "Sample",
        title = "Virus Presence Across Soybean Samples",
        subtitle = "Blue = Virus Detected | Black = Old Rotation, Red = Cullers"
    )
```

## Virus Abundance by Field Location 
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

# Data
data <- data.frame(
  Sample = c("SB1", "SB2-3", "SB5", "SB6", "SB7-8", "SB9", "SB10"),
  Field = c("Old Rotation", "Old Rotation", "Cullers", "Old Rotation", "Old Rotation", "Cullers", "Cullers"),
  SMV = c(1,1,0,1,1,1,1),
  TSWV = c(0,0,0,0,1,1,0),
  PEMV = c(1,0,0,0,0,0,0),
  PMV = c(0,0,0,1,0,0,0)
)

# Convert to long format
data_long <- data %>%
  pivot_longer(cols = SMV:PMV, names_to = "Virus", values_to = "Presence")

# Summarize total virus detections by field
field_summary <- data_long %>%
  group_by(Field) %>%
  summarise(Total_Virus_Detections = sum(Presence))

# Plot
ggplot(field_summary, aes(x = Field, y = Total_Virus_Detections, fill = Field)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Old Rotation" = "black", "Cullers" = "red")) +
  theme_minimal() +
  labs(
    title = "Virus Abundance by Field Location",
    x = "Field Location",
    y = "Total Virus Detections"
  ) +
  theme(legend.position = "none")
```

